---
title: 'Summary of available Northwest Fisheries Science Center survey data to support assessments of U.S. West Coast groundfish stocks'
author:
  - name: 'Chantel R. Wetzel'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98112'
  - name: 'R. Claire Rosemond'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2032 SE OSU Drive Building 955'
        city: 'Newport, OR'
        state: 'OR'
        postal-code: '97365'
date: today
lang: en 
keep-tex: true 
cover: 
pdf-engine: lualatex
format: 
  pdf: 
    documentclass: article
    number-sections: true
    template-partials:
      - 'support_files/before-body.tex'
      - 'support_files/_titlepage.tex'
    include-in-header:
      - 'support_files/in-header.tex'
      - 'support_files/figures.tex'
      - 'support_files/pfmc.tex'
    toc: false
    lof: false
    lot: false
    titlepage-geometry: 
      - top=2in
      - bottom=1in
      - right=1in
      - left=1in
    link-citations: true
    fig-pos: 'H'
    tab-pos: 'H'
output-file: 'data_report' 

---


```{r, results='asis'}
#| label: 'load_functions'
#| eval: true
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(table.glue)
library(glue)
library(nwfscSurvey)
library(gt)
library(gtsummary)
library(sa4ss)

firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

outer_dir <- here::here()

source(file.path(outer_dir, "R", "get_species_data_survey.R"))
source(file.path(outer_dir, "R", "get_source.R"))
source(file.path(outer_dir, "report", "ongoing_research.R"))
source(file.path(outer_dir, "report", "maturity_parameters_text.R"))

assess_data <- read.csv(file.path(here::here(), "data-processed", "2026", "assess_year_ssc_rec.csv"))
assess_data_df <- assess_data |>
    dplyr::mutate(
      species = dplyr::case_when(
        Species == "lingcod" ~ "lingcod north",
        Species == "yellowtail rockfish" ~ "yellowtail rockfish north",
        .default = Species
      )
    )
assess_data_df <- dplyr::bind_rows(
    assess_data_df,
    assess_data_df[which(assess_data_df == "lingcod"), ] |>
      dplyr::mutate(species = "lingcod south"),
    assess_data_df[which(assess_data_df == "yellowtail rockfish"), ] |>
      dplyr::mutate(
        species = "yellowtail rockfish south",
        Last_Assess = NA,
        year = NA,
        type = NA,
        SSC_Rec = NA
      )
  )

load(file.path(here::here(), "data-processed", "2026", "data_total_by_year.Rdata"))
data <- data_total_by_year |>
  dplyr::filter(Year < 2026)
# Need to make dynamic based on data file
startyr <- min(data$Year)
endyr <- max(data$Year)
assess_year <- 2027


species_data <- get_species_data_survey()
data <- dplyr::left_join(data, species_data, by = "Common_name")
n_species <- length(unique(data$Common_name))

maturity <- read.csv(here::here("data-processed", "2026", "maturity_reads_updated_2025.csv"))

``` 


{{< pagebreak >}} 

Please cite this publication as: 

Wetzel, C.R., and R.C. Rosemond. 2026. Summary of available Northwest Fisheries Science Center survey data to support assessments of U.S. West Coast groundfish stocks. Pacific Fisheries Management Council, Portland, Oregon.


{{< pagebreak >}} 

# Disclaimer {.unnumbered .unlisted}

These materials do not constitute a formal publication and are for information only. They are in a pre-review, pre-decisional state and should not be formally cited or reproduced. They are to be considered provisional and do not represent any determination or policy of NOAA or the Department of Commerce.

{{< pagebreak >}}

\hypersetup{linkcolor=.}
\setcounter{tocdepth}{3}
\tableofcontents

{{< pagebreak >}} 

\setlength{\parskip}{5mm plus1mm minus1mm}
\pagenumbering{arabic}
\setcounter{page}{1}
\setcounter{section}{0}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}

```{r, results='asis'}
#| label: 'introduction'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('01_introduction.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

{{< pagebreak >}} 

# Preliminary species identified for assessment in 2027

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "black rockfish"
)
```

{{< include 02_priority_species_report.qmd >}} 

{{< pagebreak >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "bocaccio"
)
```

{{< include 02_priority_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "lingcod north"
)
```

{{< include 02_priority_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "lingcod south"
)
```

{{< include 02_priority_species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "longspine thornyhead"
)
```

{{< include 02_priority_species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific spiny dogfish"
)
```

{{< include 02_priority_species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "petrale sole"
)
```

{{< include 02_priority_species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "redbanded rockfish"
)
```

{{< include 02_priority_species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "shortspine thornyhead"
)
```

{{< include 02_priority_species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "widow rockfish"
)
```

{{< include 02_priority_species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yellowtail rockfish south"
)
```

{{< include 02_priority_species_report.qmd >}}

{{< pagebreak >}} 

# Other groundfish species

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "arrowtooth flounder"
)
```

{{< include 03_other_species_report.qmd >}}  

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "aurora rockfish"
)
```

{{< include 03_other_species_report.qmd >}}  

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "bank rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "big skate"
)
```

{{< include 03_other_species_report.qmd >}}  

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "blackgill rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
#params <- list(
#  Common_name = "cabezon"
#)
```

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "California scorpionfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "canary rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "chilipepper"
)
```

{{< include 03_other_species_report.qmd >}}  

```{r, echo = FALSE}
# Define parameters for the first section
#params <- list(
#  Common_name = "copper rockfish"
#)
```

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "cowcod"
)
```

{{< include 03_other_species_report.qmd >}}  

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "darkblotched rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Dover sole"
)
```

{{< include 03_other_species_report.qmd >}}  

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "English sole"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "flathead sole"
)
```

{{< include 03_other_species_report.qmd >}}  

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "greenspotted rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "greenstriped rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "kelp greenling"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "longnose skate"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific cod"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific ocean perch"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific sanddab"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "quillback rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "redstripe rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "rex sole"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "rosethorn rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "rougheye and blackspotted rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "sablefish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "sharpchin rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
#params <- list(
#  Common_name = "shortraker rockfish"
#)
```

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "silvergray rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "splitnose rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "squarespot rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "starry rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "stripetail rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "vermilion and sunset rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yelloweye rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yellowmouth rockfish"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yellowtail rockfish north"
)
```

{{< include 03_other_species_report.qmd >}} 

```{r, results='asis'}
#| label: '04_survey_settings'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04_survey_settings.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: '05_acknowledgments'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('05_acknowledgments.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

