---
title: 'Summary of available Northwest Fishery Science Center survey data to support assessments of U.S. West Coast groundfish stocks'
author:
  - name: 'Chantel R. Wetzel'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98112'
  - name: 'Claire Rosemond'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2032 SE OSU Drive Building 955'
        city: 'Newport, OR'
        state: 'OR'
        postal-code: '97365'
date: today
lang: en 
keep-tex: true 
cover: 
pdf-engine: lualatex
format: 
  pdf: 
    documentclass: article
    number-sections: true
    template-partials:
      - 'support_files/before-body.tex'
      - 'support_files/_titlepage.tex'
    include-in-header:
      - 'support_files/in-header.tex'
      - 'support_files/figures.tex'
      - 'support_files/pfmc.tex'
    toc: false
    lof: false
    lot: false
    titlepage-geometry: 
      - top=2in
      - bottom=1in
      - right=1in
      - left=1in
    link-citations: true
    fig-pos: 'H'
    tab-pos: 'H'
output-file: 'data_report' 

---


```{r, results='asis'}
#| label: 'load_functions'
#| eval: true
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(table.glue)
library(glue)
library(nwfscSurvey)
library(gt)
library(gtsummary)
library(sa4ss)

firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

outer_dir <- here::here()

source(file.path(outer_dir, "R", "get_species_data_survey.R"))
source(file.path(outer_dir, "R", "get_source.R"))
source(file.path(outer_dir, "report", "ongoing_research.R"))

assess_data <- read.csv(file.path(here::here(), "data-raw", "2026", "assess_year_ssc_rec.csv"))
assess_data_df <- assess_data |>
    dplyr::mutate(
      species = dplyr::case_when(
        Species == "lingcod" ~ "lingcod north",
        Species == "yellowtail rockfish" ~ "yellowtail rockfish north",
        .default = Species
      )
    )
assess_data_df <- dplyr::bind_rows(
    assess_data_df,
    assess_data_df[which(assess_data_df == "lingcod"), ] |>
      dplyr::mutate(species = "lingcod south"),
    assess_data_df[which(assess_data_df == "yellowtail rockfish"), ] |>
      dplyr::mutate(
        species = "yellowtail rockfish south",
        Last_Assess = NA,
        year = NA,
        type = NA,
        SSC_Rec = NA
      )
  )

load(file.path(here::here(), "data-processed", "2026", "data_total_by_year.Rdata"))
data <- data_total_by_year
# Need to make dynamic based on data file
startyr <- min(data$Year)
endyr <- max(data$Year)
assess_year <- 2027

species_data <- get_species_data_survey()
data <- dplyr::left_join(data, species_data, by = "Common_name")
n_species <- length(unique(data$Common_name))

maturity <- read.csv(here::here("data-processed", "2026", "maturity_reads_updated_2025.csv"))

``` 


{{< pagebreak >}} 

Please cite this publication as: 

Wetzel, C.R., and C. Rosemond. 2026. Summary of available Northwest Fishery Science Center survey data to support assessments of U.S. West Coast groundfish stocks. Pacific Fisheries Management Council, Portland, Oregon.


{{< pagebreak >}} 

\hypersetup{linkcolor=.}
\setcounter{tocdepth}{3}
\tableofcontents

{{< pagebreak >}} 

\setlength{\parskip}{5mm plus1mm minus1mm}
\pagenumbering{arabic}
\setcounter{page}{1}
\setcounter{section}{0}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}

```{r, results='asis'}
#| label: 'introduction'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('01_introduction.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

{{< pagebreak >}} 

# Groundfish species observed by surveys conducted by the NWFSC


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "arrowtooth flounder"
)
```

{{< include species_report.qmd >}} 

{{< pagebreak >}} 


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "aurora rockfish"
)
```

{{< include species_report.qmd >}} 


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "bank rockfish"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "big skate"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
#params <- list(
#  Common_name = "black rockfish"
#)
```


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "blackgill rockfish"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "bocaccio"
)
```

{{< include species_report.qmd >}} 


```{r, echo = FALSE}
# Define parameters for the first section
#params <- list(
#  Common_name = "cabezon"
#)
```


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "California scorpionfish"
)
```

{{< include species_report.qmd >}} 


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "canary rockfish"
)
```

{{< include species_report.qmd >}} 


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "chilipepper"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
#params <- list(
#  Common_name = "copper rockfish"
#)
```

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "darkblotched rockfish"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Dover sole"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "English sole"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "flathead sole"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "greenspotted rockfish"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "greenstriped rockfish"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "kelp greenling"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "lingcod north"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "lingcod south"
)
```

{{< include species_report.qmd >}} 

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "longnose skate"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "longspine thornyhead"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific cod"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific ocean perch"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific sanddab"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "Pacific spiny dogfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "petrale sole"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "quillback rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "redbanded rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "redstripe rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "rex sole"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "rosethorn rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "rougheye and blackspotted rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "sablefish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "sharpchin rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
#params <- list(
#  Common_name = "shortraker rockfish"
#)
```


```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "shortspine thornyhead"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "silvergray rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "splitnose rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "squarespot rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "starry rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "stripetail rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "vermilion and sunset rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "widow rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yelloweye rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yellowmouth rockfish"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yellowtail rockfish north"
)
```

{{< include species_report.qmd >}}

```{r, echo = FALSE}
# Define parameters for the first section
params <- list(
  Common_name = "yellowtail rockfish south"
)
```

{{< include species_report.qmd >}}



```{r, results='asis'}
#| label: '03_survey_settings'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('03_survey_settings.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: '04_acknowledgments'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04_acknowledgments.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

